Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-14 paper/sec-analysis.tex
more from Daniela's comments

Now we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in a high pile-up (PU) environment.
%
To study the impact of PU,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
the PU subtraction.

%
In order to subtract PU in hadronic collisions, a number of techniques
are available~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.\footnote{
These techniques have also important applications in the subtraction
of the UE/MPI contamination for jet reconstruction
in heavy ion collisions~\cite{Cacciari:2010te}.
}
%
In this work, PU  will be subtracted by means
of the {\tt SoftKiller} (SK)
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet},
whose performance has been shown to
improve the commonly used area-based subtraction~\cite{Cacciari:2009dp}.
%
The idea underlying {\tt SoftKiller} is based on eliminating particles
below a given cut-off in their transverse momentum, $p_T^{\rm (cut)}$, whose
value is dynamically determined in a way that makes the event-wide
transverse-momentum flow density $\rho$ vanish.
%
This $p_T$ flow density is defined as
\be
\rho\equiv{\rm median}_i \Bigg\{ \frac{p_{Ti}}{A_i}\Bigg\} \, ,
\ee
where the median is computed over all the patches $i$ with area
$A_i$ and transverse momentum $p_{Ti}$ in which the $\lp \eta,\phi\rp$ plane
is partitioned.
%
From its definition in terms of the median,
it follows that the value of $p_T^{(\rm cut)}$
will be dynamically raised until half of the patches have
$p_{Ti}=0$.
%
The size and number of these patches is a free parameter of the algorithm -
here we will use square patches with length $a=0.4$.
%
We restrict ourselves to the central rapidity region,
$|\eta| \le 2.5$, for the estimation of the
$p_T$ flow density $\rho$.
%
We apply {\tt SoftKiller}
to particles at the end of the parton shower, before
jet clustering.

@@

To validate PU subtraction,
we now compare representative kinematic distributions
in the case without PU and in the case
with PU subtracted with {\tt SoftKiller}.
%
We compare first
in Fig.~\ref{fig:PUvalidation} the
invariant mass distributions of Higgs candidates for signal
events in the resolved category.
%
We show three curves: without PU,
with PU $\la n_{\rm PU}\ra=80$ but without any subtraction, and the
same but now with the SK subtraction.
%
If PU is not subtracted there is a large shift in the Higgs mass
peak, by about 40 GeV.
%
Once SK subtraction is performed, we recover a distribution much closer
to the original ones, with only a small shift of $\simeq 5$ GeV
and a broadening of the mass
distribution.
%

Next, we compare additional
 distributions with
and without PU, with $\la n_{\rm PU}\ra=80$ in the
former
case (and SK subtraction).
%
In Fig.~\ref{fig:m_H_PU} we show the invariant mass distribution
of the leading and subleading Higgs candidates,
this time corresponding
to the boosted category.
%
These distributions are plotted after the $b$-tagging, that is,
before they are used as input to the MVA.
%
As we can see, the residual effects of PU
after the {\tt SoftKiller} subtraction are smaller
in the boosted selection as compared to the resolved case,
with the position of the Higgs mass peaks essentially
unchanged, and only a moderate broadening of the
mass distribution.
%

@@

In Fig.~\ref{fig:mHH_PU}
we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in both for the boosted and
for the resolved categories.
%
In the case of the $p_T^{h1}$ distribution, the differences between the selection
criteria for the resolved
and boosted categories is reflected in the rightward shift of the latter.
%
The effect of PU is small in the boosted case, and generally moderate
in the resolved case, except for large $p_T$ values.
%
Similar observation can be derived for the case of
the di-Higgs invariant mass $m_{hh}$ distribution, similar 
apply.
%
Therefore, we can consider the PU subtraction strategy
as validated for the purposes of this study, although
there could still be room for optimisation.

@@

We can also assess the impact of PU in some of
substructure variables that will be 
used as input to the MVA in the boosted category.
%
In Fig.~\ref{fig:Substructure_PU} we show the 2-to-1 subjettiness ratio
$\tau_{21}$, Eq.~(\ref{eq:tau21}), and the ratio
of energy correlation functions, $D_2^{(\beta)}$,
Eq.~(\ref{eq:d2}), in both cases
corresponding to the leading Higgs candidate.
%
Comparing the signal distributions in the cases without PU and
with PU subtracted with SK, we observe that both substructure variables
are reasonably robust in a high-PU environment.
%
The ratio of energy correlation functions $D_2^{(\beta)}$ 
was indeed constructed~\cite{Larkoski:2013eya}
to be resilient
with respect to PU contamination.
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/m_h0_bst_comp_back.pdf}
  \includegraphics[width=0.49\textwidth]{plots/pt_h0_bst_comp_back.pdf}
   \includegraphics[width=0.49\textwidth]{plots/tau21_h1_bst_comp_back.pdf}
  \includegraphics[width=0.49\textwidth]{plots/split12_h0_bst_comp_back.pdf}
   \caption{\small
     Comparison of kinematic distributions, in
     the boosted category, for signal and background events
     with and without PU: the invariant mass,  $p_T$,
     and the substructure variables $\tau_{21}$ and $\sqrt{d_{12}}$
    for the leading Higgs candidate.
     %
 }
\label{fig:signal-vs-back-boosted}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Here we turn to discuss how the description of kinematic
distributions for $hh\to 4b$ signal
and background processes are
modified in a high pile-up (PU) environment.
%
To study the impact of PU,
Minimum Bias (MB) events,
including Multiple Parton Interactions (MPI), have been generated
with {\tt Pythia8}, and then
superimposed to the signal
and background samples described in Sect.~\ref{mcgeneration}.
%
We have explored two scenarios,
one with a mean number of
PU vertices per bunch crossing of $\la n_{\rm PU}\ra=80$, and another
with $\la n_{\rm PU}\ra=150$.
%
While the latter is closer to the HL-LHC forecasts,
we adopt $\la n_{\rm PU}\ra=80$ as our baseline,
given that it
is beyond the scope of this paper to fully optimize
the PU subtraction.

%
In order to subtract PU in hadronic collisions, a number of techniques
are available~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.\footnote{
These techniques have also important applications in the subtraction
of the UE/MPI contamination for jet reconstruction
in heavy ion collisions~\cite{Cacciari:2010te}.
}
%
In this work, PU  is subtracted
with the {\tt SoftKiller} (SK)
method~\cite{Cacciari:2014gra}, as implemented in {\tt FastJet},
whose performance has been shown to
improve the commonly used area-based subtraction~\cite{Cacciari:2009dp}.
%
The idea underlying {\tt SoftKiller} is based on eliminating particles
below a given cut-off in their transverse momentum, $p_T^{\rm (cut)}$, whose
value is dynamically determined in a way that makes the event-wide
transverse-momentum flow density $\rho$ vanish.
%
This $p_T$ flow density is defined as
\be
\rho\equiv{\rm median}_i \Bigg\{ \frac{p_{Ti}}{A_i}\Bigg\} \, ,
\ee
where the median is computed over all the regions $i$ with area
$A_i$ and transverse momentum $p_{Ti}$ in which the $\lp \eta,\phi\rp$ plane
is partitioned.
%
From its definition in terms of the median,
it follows that the value of $p_T^{(\rm cut)}$
will be dynamically raised until half of the regions have
$p_{Ti}=0$.
%
The size and number of these regions is a free parameter of the algorithm -
here we will use square regions with length $a=0.4$.
%
We restrict ourselves to the central rapidity region,
$|\eta| \le 2.5$, for the estimation of the
$p_T$ flow density $\rho$.
%
We apply {\tt SoftKiller}
to particles at the end of the parton shower, before
jet clustering.

@@

To validate PU subtraction,
we compare representative kinematic distributions
without PU, with PU, and with PU after SK subtraction.
%
In Fig.~\ref{fig:PUvalidation} we show the
invariant mass distributions of Higgs candidates for signal
events in the resolved category.
%
If PU is not subtracted there is a large shift in the Higgs mass
peak, by about 40 GeV.
%
Once SK subtraction is performed, we recover a distribution much closer
to the ones without PU, with only a small shift of $\simeq 5$ GeV
and a broadening of the mass
distribution.
%

In Fig.~\ref{fig:m_H_PU} we show the invariant mass distribution
of the leading and subleading Higgs candidates,
this time corresponding
to the boosted category.
%
These distributions are plotted after
$b$-tagging, that is,
before they are used as input to the MVA.
%
The residual effects of PU
after the {\tt SoftKiller} subtraction are smaller
in the boosted selection as compared to the resolved case,
with the position of the Higgs mass peak essentially
unchanged, and only a moderate broadening of the
mass distribution.
%

@@

In Fig.~\ref{fig:mHH_PU}
we compare the transverse momentum of the leading Higgs
candidate, $p_t^{h_1}$ and the invariant mass of the di-Higgs system
$m_{hh}$, in both for the boosted and
for the resolved categories.
%
In the case of the $p_T^{h1}$ distribution, the differences between the selection
criteria for the resolved
and boosted categories is reflected in the rightward shift of the latter.
%
The effect of PU is small in the boosted case, and generally moderate
in the resolved case, except for large $p_T$ values.
%
A similar behaviour is observed in the di-Higgs invariant mass distribution.
%
Therefore, we can consider the PU subtraction strategy
as validated for the purposes of this study, although
further optimisation should be possible.

@@

We can also assess the impact of PU in some of
substructure variables that will be 
used as input to the MVA in the boosted category.
%
In Fig.~\ref{fig:Substructure_PU} we show the 2-to-1 subjettiness ratio
$\tau_{21}$, Eq.~(\ref{eq:tau21}), and the ratio
of energy correlation functions, $D_2^{(\beta)}$,
Eq.~(\ref{eq:d2}), for the leading Higgs candidate.
%
Comparing the signal distributions in the cases without PU and
with PU subtracted with SK, we observe that both substructure variables
are reasonably robust in a high-PU environment.
%
The ratio of energy correlation functions $D_2^{(\beta)}$ 
was indeed constructed~\cite{Larkoski:2013eya}
to be robust against PU contamination.
%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
  \includegraphics[width=0.49\textwidth]{plots/pt_h0_bst_comp_back.pdf}
   \includegraphics[width=0.49\textwidth]{plots/tau21_h1_bst_comp_back.pdf}
   \includegraphics[width=0.49\textwidth]{plots/split12_h0_bst_comp_back.pdf}
   \includegraphics[width=0.49\textwidth]{plots/m_h0_bst_comp_back.pdf}
   \caption{\small
     Comparison of kinematic distributions, in
     the boosted category, for signal and background events
    in the case of PU subtracted with SK:
     the transverse momentum  $p_T$ and
     and the substructure variables $\tau_{21}$ and $\sqrt{d_{12}}$
     for the leading Higgs candidate, and the invariant
     mass of the di-Higgs system $m_{hh}$.
     %
 }
\label{fig:signal-vs-back-boosted}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
